### 2018-06-13 output/changes-list/vaccines.Rmd
added additional references and discussion of meta-analyses

---
title: "Vaccines"
author: "Sam Abbott"
date: "13/06/2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Text

Highlighted the novel findings over the Abubakar et al. 2013 (10.3310/hta17370) review with the following in the discussion;

- The meta-analysis performed by Abubakar et al. also found consistent evidence for this association but was limited by the number, and quality, of studies available.[@Abubakar2013] In contrast to our study, both of these meta-analyses estimated the protection from TB mortality in BCG vaccinated individuals rather than in BCG vaccinated cases. Additionally, neither study explored the association between BCG vaccination and all-cause mortality or recurrent TB.

# Highlights

Modified previous summary bullets to be highlights added the following sentance to the second bullet;

- There was little evidence of an association with other TB outcomes.

# Keywords

Added the following;

- non-specific
- mortality

# References

I have added 4 new references to the introduction;

- BCG vaccine: WHO position paper (https://doi.org/10.1016/j.vaccine.2018.03.009)
- Causes of variation in BCG vaccine efficacy: Examining evidence from the BCG REVAC cluster randomized trial to explore the masking and the blocking hypotheses (https://doi.org/10.1016/j.vaccine.2014.05.042)
- BCG scar and positive tuberculin reaction associated with reduced child mortality in West Africa: A non-specific beneficial effect of BCG? (https://doi.org/10.1016/S0264-410X(03)00181-6)