### 2018-06-13 output/changes-list/vaccines.Rmd
packrat updates, added new version DOI

---
title: "Vaccines"
author: "Sam Abbott"
date: "13/06/2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Text

Highlighted the novel findings over the Abubakar et al. 2013 (10.3310/hta17370) review with the following in the discussion;

- The meta-analysis performed by Abubakar et al. also found consistent evidence for this association but was limited by the number, and quality, of studies available.[@Abubakar2013] In contrast to our study, both of these meta-analyses estimated the protection from TB mortality in BCG vaccinated individuals rather than in BCG vaccinated cases who had died from any cause. Additionally, neither study explored the association between BCG vaccination and all-cause mortality or recurrent TB.

# Highlights

Modified previous summary bullets to be highlights added the following sentance to the second bullet;

- There was little evidence of an association with other TB outcomes.

# Keywords

Added the following;

- non-specific
- mortality

# References

I have added 3 new references to the introduction;

- BCG vaccine: WHO position paper (https://doi.org/10.1016/j.vaccine.2018.03.009)
- Causes of variation in BCG vaccine efficacy: Examining evidence from the BCG REVAC cluster randomized trial to explore the masking and the blocking hypotheses (https://doi.org/10.1016/j.vaccine.2014.05.042)
- BCG scar and positive tuberculin reaction associated with reduced child mortality in West Africa: A non-specific beneficial effect of BCG? (https://doi.org/10.1016/S0264-410X(03)00181-6)