### 2018-06-13 packrat/lib/x86_64-apple-darwin15.6.0/3.5.0/PKPDmodels/doc/UsersGuide.Rnw

\documentclass[12pt]{article}
\usepackage{Sweave,amsmath,amsfonts,bm}
\usepackage[authoryear,round]{natbib}
\bibliographystyle{plainnat}
\DefineVerbatimEnvironment{Sinput}{Verbatim}
{formatcom={\vspace{-1ex}},fontshape=sl,
  fontfamily=courier,fontseries=b, fontsize=\footnotesize}
\DefineVerbatimEnvironment{Soutput}{Verbatim}
{formatcom={\vspace{-1ex}},fontfamily=courier,fontseries=b,%
  fontsize=\footnotesize}
%%\VignetteIndexEntry{User's Guide}
%%\VignetteDepends{PKPDmodels,lattice}
\title{User's Guide for the PKPDmodels Package}
\author{
  Douglas Bates\\
  Dept. of Statistics, University of Wisconsin - Madison, U.S.A.\\
  and\\
  France Mentr\'{e}\\
  UMR738, INSERM, University Paris Diderot\\
}
\SweaveOpts{engine=R,eps=FALSE,pdf=TRUE,strip.white=true,keep.source=TRUE}
\SweaveOpts{include=TRUE}
\setkeys{Gin}{width=\textwidth}
%% commands
\makeatletter
\newcommand\code{\bgroup\@makeother\_\@makeother\~\@makeother\$\@codex}
\def\@codex#1{{\normalfont\ttfamily\hyphenchar\font=-1 #1}\egroup}
\makeatother
%%\let\code=\texttt
\let\proglang=\textsf
\newcommand{\pkg}[1]{{\fontseries{b}\selectfont #1}}
\newcommand{\Cl}{\ensuremath{\mathit{Cl}}}
\begin{document}
<<preliminaries,echo=FALSE,results=hide>>=
options(width=69,digits=5,show.signif.stars=FALSE,str=strOptions(strict.width="cut"))
library(PKPDmodels)
library(lattice)
#lattice.options(default.theme = function() standard.theme())
lattice.options(default.theme = function() standard.theme(color=FALSE))
@
\maketitle
\begin{abstract}
  The \pkg{PKPDmodels} package provides a function, \code{PKmod},
  that is used to generate model functions for 1, 2, or 3 compartment
  pharmacokinetic models with linear elimination kinetics.  The form
  of administration can be intravenous bolus, infusion or oral
  administration with first-order absorption.  The model can be for
  single dose, multiple doses or steady-state conditions.
\end{abstract}

\section{Introduction}
\label{sec:Intro}

A pharmacokinetic model describes the concentration of an analyte in
the central compartment of a compartment model as a function of time
and other covariates such as dose.  The \code{PKmod} function
generates model functions for pharmacokinetic models with linear
elimination kinetics.  The mode of administration can be intravenous
bolus, continuous infusion or oral with first-order absorption
kinetics.  The dosage can be a single dose, multiple doses or the
steady-state condition after multiple doses.

The pharmacokinetic parameters can be expressed as an elimination rate
constant, $k$ (units of 1/time), the clearance rate, $\Cl$ (units of
volume per unit time) or volume of distribution, $V$ (units of
volume).  Only two of these three parameters are specified because of
the relationship
\begin{displaymath}
  k = \frac{\Cl}{V}
\end{displaymath}

Some types of models include other parameters such as the absorption
rate constant, $k_a$, for oral doses with first-order absorption.

By default the function that is returned is a byte-compiled function
that evaluates both the expected values of the concentration in the
central compartment and the gradient with respect to the parameters.
There is an option to create a function that will evaluate the
expected value of the response and the gradient and the Hessian.
Although a
default set of parameters is provided, these can be modified by a
set of parameter transformations.  For example, the default
parameterization for a one-compartment model uses the
elimination rate constant, $k$, and the clearance, $\Cl$ but the model
can be expressed in terms of the volume of distribution, $V$, and the
clearance, $\Cl$, by including the transformation \code{k ~ V/Cl}.
For population pharmacokinetics it is often helpful to transform from
parameters such as $\Cl$ and $V$ to their logarithms.

\section{A simple example}
\label{sec:simple}

A PK model with linear elimination for a single-dose bolus injection
can be expressed in terms of the elimination rate constant, $k$, and
the volume of distribution, $V$, as
\begin{displaymath}
  C(t) = \frac{D\exp(-k\,t)}{V}
\end{displaymath}
where $C(t)$ is the concentration in the central compartment at time
$t$ (since administration of the dose) and $D$ is the dose.  The
corresponding formula in \proglang{R} is 
<<bolus_1cpt_sd_Vk_expr>>=
PKexpr("bolus", "sd")
@

If we wish to transform to another set of parameters, we specify the
transformations as a list of formulas.  For example, to express the
model in terms of the volume of distribution, $V$, and the clearance,
$\Cl$,
<<bolus1cptSdVClExpr>>=
PKexpr("bolus", "sd", list(k ~ Cl/V))
@ 
or, to express the model in terms of $\log(V)$ and $\log(\Cl)$,
<<bolus1cptSdlVlClExpr>>=
PKexpr("bolus", "sd", list(k ~ Cl/V, Cl ~ exp(lCl), V ~ exp(lV)))
@
Note that a substitution formula has the parameter name on the
left and the expression to be substituted for that parameter on the
right. When there are multiple substitution formulas they are
evaluated left to right.

The function \code{PK1expr} exists solely so that the user can verify
that the formula being used is what they expect.  In typical usage we
create the model function directly.
<<bolus_1cpt_sd_Vk_expr>>=
(bolus1cptSdVk <- PKmod("bolus", "sd"))
@ 
When this function is evaluated at given values of \code{dose},
\code{time} and the pharmacokinetic parameters, it returns the
function evaluations with an attribute called \code{"gradient"}.
<<bolusSdVkEval>>=
tvals <- 0:10
bolus1cptSdVk(dose=1, t=tvals, k=0.3, V=1)
@

When fitting pharmacokinetic models, especially population
pharmacokinetic models, it is helpful to have this analytic gradient
evaluation rather than calculating numerical gradients.  Because
common sub-expressions in the gradient columns and the model function
are evaluated once only and because the resulting function consists of
rather simple byte-compiled expressions the evaluation of the model
function and gradient is essentially as fast as the evaluation of the
model function itself.

The evaluation of common sub-expressions is more obvious when several
parameter transformations are applied
<<bolusSdlVlCl>>=
(bolus1cptSdlVlCl <-
 PKmod("bolus", "sd", list(k ~ Cl/V, Cl ~ exp(lCl), V ~ exp(lV))))
@
where we can see that the clearance, $\Cl=\exp(\mathit{lCl})$
(\code{.expr1}), the volume of distribution, $V=\exp{\mathit(lV)}$
(\code{.expr2}), and the elimination rate constant, $k=\Cl/V$
(\code{.expr3}) are each evaluated once only.

Expressing the parameters in this formulation we obtain the same
function evaluation as before
<<bolusSdlVlClEval>>=
bolus1cptSdlVlCl(dose=1, t=tvals, lCl=log(0.3), lV=0)
@
Naturally the gradient has changed because we are calculating the
derivatives with respect to a different set of parameters.

\section{Single oral dose}
\label{sec:simple}

In phase 1 and phase 2 clinical trials population pharmacokinetics are
often assessed by sampling at many closely-spaced time points in a
small number of subjects after a single oral dose.  The \code{Theoph}
data set in the \pkg{datasets} package comes from such a study of the
drug theophylline which is used in the treatment of asthma.
<<strTheoph, echo=FALSE>>=
str(Theoph <- data.frame(lapply(Theoph, function(x) x)))
@ 
These data are shown in Figure~\ref{fig:Theophplot}.
\begin{figure}[tb]
  \centering
<<Theophplot,fig=TRUE,echo=FALSE,height=3>>=
print(xyplot(conc ~ Time|reorder(Subject, Theoph$conc, max),
             Theoph, type=c("g","b"),
             xlab="Time since drug administration (hr)",
             ylab="Serum concentration (mg/l)",
             aspect="xy", layout=c(6,2)))
@   
\caption{Concentration of theophylline versus time for 12 subjects
  following a single oral dose of the drug.  The panels have been
  ordered (left to right starting at the bottom row) by increasing
  maximum concentration.}
  \label{fig:Theophplot}
\end{figure}

A model function for a single oral dose with first-order absorption
using the parameters $\log(k_a)$, $\log(V)$ and $\log(\Cl)$ is
<<oral1cptSdlkalVlCl>>=
oral1cptSdlkalVlCl <-
 PKmod("oral", "sd", list(ka ~ exp(lka), k ~ exp(lCl)/V, V ~ exp(lV)))
@ 
corresponding to the formula
<<oral1cptSdlkalVlCl>>=
PKexpr("oral", "sd", list(ka ~ exp(lka), k ~ exp(lCl)/V, V ~ exp(lV)))
@

Initial values for the parameters are taken to be $\log(V)_0=-1$,
$\log(k_a)_0=0.5$ and $\log(\Cl)_0=-4$ producing a fitted model
<<fm1>>=
summary(fm1 <- nls(conc ~ oral1cptSdlkalVlCl(Dose, Time, lV, lka, lCl),
                   Theoph, start=c(lV=-1, lka=0.5, lCl=-4),
                   subset=Subject==1), corr=TRUE)
@

\end{document}